--- 
title: "My project"
# author: "Tiana Chen"
date: "2024-10-24"
output: "github_document"
---

# load packages and dataset
```{r}
# Set a CRAN mirror before installing packages
options(repos = c(CRAN = "https://cloud.r-project.org"))

# Install and load the moments package
install.packages("moments")
library(moments)
library(haven) #to download the package
library(dplyr) #re-code variable
library(ggplot2)
library(bruceR)
# Install and load car package if not already installed
install.packages("car")
library(car)

#hope it's the new one, yes it is 
load ("/Users/chenxuejing/Downloads/ICPSR_36850/DS0003/36850-0003-Data.rda")
dataset <- da36850.0003
rm(da36850.0003)
dataset <- dataset %>%
  select(PSYDIST2, SLET2, DEMGEN2, DEMRA2013, KESPDS201, KESPDS202, KESPDS203, KESPDS204, KESPDS205, KESPDS206, SLE01, SLE02, SLE03, SLE04, SLE05, SLE06, SLE07, SLE08, SLE09, SLE10, SLE11, SLE12, SLE13, DEMRA2013)
```

```{r}
#check normality assumption
# Histogram
ggplot(dataset, aes(x = PSYDIST2)) +
  geom_histogram(binwidth = 0.5, fill = 'blue', color = 'black') +
  ggtitle('Histogram of PSYDIST2') +
  xlab('PSYDIST2') +
  ylab('Frequency')

#check equal variance assumption by using Levene's Test for equal variance
leveneTest(PSYDIST2 ~ factor(SLET2), data = dataset)
ggplot(dataset, aes(x = factor(SLET2), y = PSYDIST2)) +
  geom_boxplot(fill = 'lightblue') +
  ggtitle('Boxplot of PSYDIST2 by SLET2') +
  xlab('SLET2') +
  ylab('PSYDIST2')

dataset$PSYDIST2_log <- log(dataset$PSYDIST2)

#check normality assumptions for psydist2_log
skewness(dataset$PSYDIST2_log, na.rm = TRUE)
kurtosis(dataset$PSYDIST2_log, na.rm = TRUE)

ggplot(dataset, aes(x = factor(SLET2), y = PSYDIST2_log)) +
  geom_boxplot(fill = 'lightblue') +
  ggtitle('Boxplot of PSYDIST2_log by SLET2') +
  xlab('SLET2') +
  ylab('PSYDIST2')
```

```{r}
#lab8 
# run the substantive analysis to test my research question. (Correlation model with gender interaction)
Corr(dataset)

# there is a significant correlation between my two IVs. Multicollinearity occurs. Try to solve it by standardizing my variables:
# Standardize the 'stressful events' variable (SLET2)
dataset$SLET2_standardized <- scale(dataset$SLET2)

# Convert gender variable (DEMGEN2) to numeric, 0 = male; 1 = female, then standardize it
dataset$DEMGEN2 <- as.numeric(dataset$DEMGEN2)
dataset$DEMGEN2 <- ifelse(dataset$DEMGEN2 == 2, 1, 0)

# Check the result, 332 males and 401 females, then standardize it
table(dataset$DEMGEN2)
dataset$DEMGEN2_standardized <- scale(dataset$DEMGEN2)

# standardize dependent variable (PSYDIST2) if needed 
dataset$PSYDIST2_standardized <- scale(dataset$PSYDIST2)

# View the standardized data
head(data)

#rename the standardized dataset as "data1"
data1 <- dataset %>% 
  select(SLET2_standardized, DEMGEN2_standardized, PSYDIST2_standardized)
Corr(data1)

#moderation:
PROCESS(dataset, y = "PSYDIST2", x = "SLET2", mods = c("DEMGEN2"))

#check if there's a missing value of gender, and there's one row that has missing values 
sum(is.na(dataset$DEMGEN2))
dataset <- dataset[!is.na(dataset$DEMGEN2), ]
```


```{r}
model<-lm(PSYDIST2 ~ SLET2 + DEMGEN2 + DEMRA2013, data = dataset)

#step 2: check the assumptions
check_model(model)
```


```{r}
ggplot(dataset, aes(x = SLET2, y = PSYDIST2)) + geom_point() + geom_smooth() + theme_bruce()

# Plot the interaction effect of stressful events and gender on psychological distress
ggplot(dataset, aes(x = SLET2, y = PSYDIST2, color = factor(DEMGEN2))) + 
  geom_smooth(method = "lm") +  facet_wrap(~DEMGEN2) +
  ggtitle("Interaction of SLET2 and DEMGEN2 on Psychological Distress (PSYDIST2)") +
  xlab("Numbers of Stressful Events (SLET2)") +
  ylab("Psychological Distress (PSYDIST2)") +
  scale_color_discrete(name = "Gender", labels = c("Male", "Female"))

```

```{r}
#explanatory research
Corr(dataset)

#Moderation:
PROCESS(dataset, y = "PSYDIST2", x = "SLET2", mods = c("DEMRA2013"))

#visualize my data
ggplot(dataset, aes(x = SLET2, y = PSYDIST2)) + geom_point() + geom_smooth() + theme_bruce()

# Plot the interaction effect of stressful events and gender on psychological distress
ggplot(dataset, aes(x = SLET2, y = PSYDIST2, color = factor(DEMRA2013))) + 
  geom_smooth(method = "lm") +  facet_wrap(~DEMRA2013) +
  ggtitle("Interaction of Stressful Events (SLET2) and Asian ethinicity (DEMRA2013) on Psychological Distress (PSYDIST2)") +
  xlab("Stressful Events (SLET2)") +
  ylab("Psychological Distress (PSYDIST2)") +
  scale_color_discrete(name = "Asian ethinicity", labels = c("Non-Asian", "Asian"))

```
# Test the DV: Non-psychological distress scale (PSYDIST2: Mean of KESPDS201-KESPDS206)
# Because this variable has 6 items, we shall run the reliability and factor analysis.
```{r}
#Convert items (KESPDS201-KESPDS206) to numeric: 1 = none of the time, 2 = a little of the time, 3 = some of the time, 4 = most of the time, 5 = all of the time
# if there's a reversed-scored item (no matter if you recode or not) its actual value doesn't change 
dataset$KESPDS201 <- as.numeric(dataset$KESPDS201)
dataset$KESPDS202 <- as.numeric(dataset$KESPDS202)
dataset$KESPDS203 <- as.numeric(dataset$KESPDS203)
dataset$KESPDS204 <- as.numeric(dataset$KESPDS204)
dataset$KESPDS205 <- as.numeric(dataset$KESPDS205)
dataset$KESPDS206 <- as.numeric(dataset$KESPDS206)

#test the internal consistency reliability for DV [non-psychological distress scale (PSYDIST2: Mean of KESPDS201-KESPDS206)] by looking at the Cronbach's alpha (should be >0.9, but if it is larger than 0.95, it means there are redundant items)
Alpha(dataset, "KESPDS", 201:206)

# test how many factors we have are identified in this study (E-value, and Scree plot), if there're two factors, it means some items are cross-loading. For minimizing item cross-loading and improve the clarity , use rotation (make the difference more polarized) Then use the othrogonal or oblique assumption

# EFA analysis: look at the factor loading of one factor (0-1) and if it's below 0.4 is not acceptable.
EFA(dataset, "KESPDS", 201:206, method = "pa", plot.scree = TRUE, nfactors = c("parallel"))
```

# Test the IV: Stressful events (SLET2: The sum of SLE01-SLE13) 
# Even this variable has 13 items, all the stressful life events items are binary rather than on a scale of 1-5 or 1-7. So it is unable to run cronbach's alpha or EFA on a measure like this.
